\documentclass[11pt]{article}
\usepackage{acl}
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{booktabs}

\title{HedgeBERT: Hedge-Aware Sentiment Classification via CLS Feature Injection}

\author{
  Minh Tran \\
  Wichita State University \\
  \texttt{K699E569@wichita.edu}
  \And
  An Nguyen \\
  Wichita State University \\
  \texttt{R374G648@wichita.edu}
}

\begin{document}
\maketitle

\begin{abstract}
Sentiment analysis models struggle with classifying hedged language, as it can be
less direct and ambiguous---phrases such as ``I guess it is sort of decent'' carry
real but weakened sentiment rather than no sentiment at all. Hedged language is also
common in many real-world scenarios such as product reviews and customer feedback,
where mistaking a negative hedged review for a positive one can lead organizations
to misread user opinion entirely. Despite this, standard benchmarks such as SST-2
\cite{socher2013recursive} are dominated by clear-polarity examples and do not
expose this failure mode, meaning no prior work has measured the accuracy gap on
hedged versus direct sentiment as a diagnostic category. We propose HedgeBERT, a
modified BERT model that encodes a hedge intensity score---computed by a rule-based
scorer drawing on BioScope \cite{vincze2008bioscope} and the SFU Review Corpus
\cite{konstantinova2012sfu}---into the CLS token to help BERT better make decisions
on hedged language. We expect HedgeBERT to demonstrate measurable accuracy
improvement on hedged consumer sentiment without degrading performance on direct
sentiment.
\end{abstract}

\section{Research Question}

Do standard sentiment classifiers perform significantly worse on hedged sentiment
expressions than on direct ones, and can explicitly encoding hedge intensity as a
model feature close that gap? This question targets a measurable failure mode: 
hedged language is not neutral, yet models trained on polarity-skewed
benchmarks treat it as though it were. HedgeBERT is designed to test whether a
lightweight, architecture-level intervention can remedy this without resorting to
large-scale model changes.

\section{Motivation}

Hedged expressions pervade consumer language. Phrases such as ``I suppose it works
well enough'' or ``it is not terrible'' signal negative or weakly positive sentiment,
yet their syntactic indirection causes standard classifiers to misfire or abstain.
Prior work has long recognized hedging as a linguistic phenomenon
\cite{taboada2011lexicon, vincze2008bioscope}, but no study has isolated its effect
on neural sentiment accuracy as a diagnostic category.

This gap matters because sentiment analysis is deployed in high-stakes
settings---customer feedback, healthcare, and market research---where hedged language
is especially common. Standard benchmarks such as SST-2 \cite{socher2013recursive}
and IMDb are dominated by clear-polarity examples, so a model that simply learns
strong cue words passes existing evaluations while failing silently on hedged input.
Measuring and closing this gap is therefore both a scientific and a practical
necessity.

\section{Related Work and Novelty}

Sentiment analysis has a rich history. \citet{pang2008opinion} survey the field and
establish the assumption that polarity is recoverable from lexical cues, an
assumption that breaks down under hedging. \citet{socher2013recursive} introduce the
Stanford Sentiment Treebank, which, despite its phrase-level annotations, remains
dominated by direct-polarity examples. \citet{taboada2011lexicon} recognize hedging
as a problem in their SO-CAL system---blocking polarity propagation across irrealis
operators---but do not measure or correct for it in a neural setting.

Hedge cue taxonomies come from two corpora. \citet{vincze2008bioscope} introduce
BioScope, annotating scientific text for speculation and negation.
\citet{konstantinova2012sfu} extend this to consumer reviews in the SFU Review
Corpus, providing the speculation annotations that HedgeBERT uses to construct its
benchmark. Neither corpus has been crossed with sentiment labels for diagnostic
evaluation. On the model side, \citet{devlin2019bert} establish BERT and its
CLS-token representation as the standard injection point for sentence-level features,
while \citet{sanh2019distilbert} offer DistilBERT as a lighter alternative. The
closest architectural precedent is \citet{fasulo2025checkthat}, who inject
subjectivity signals at the CLS level for a related task; HedgeBERT applies the same
mechanism in the inverse direction.

HedgeBERT's novelty is threefold: (1) it constructs the first hedged-sentiment
diagnostic benchmark by crossing SFU speculation annotations with star-rating ground
truth; (2) it measures the accuracy gap between hedged and direct sentiment as an
explicit experimental quantity; and (3) it addresses that gap by injecting a hedge
intensity scalar at the CLS level, a targeted and interpretable intervention not
previously applied to sentiment.

\section{Proposed Methodology}

\paragraph{Hedged-Sentiment Benchmark.} The benchmark is drawn from the SFU Review
Corpus \cite{konstantinova2012sfu}. Sentences are partitioned into a hedged subset
(carrying at least one SFU speculation annotation) and a direct subset (zero
annotations). Ground truth is derived from star ratings: 1--2 stars are negative,
4--5 stars are positive, and 3-star examples are excluded. The target is
approximately 250 balanced examples matched by sentiment class and domain. This
benchmark is held out entirely and is never used for training.

\paragraph{Hedge Intensity Scorer.} A rule-based scorer maps a raw sentence to a
float in $[0.0, 1.0]$. The cue inventory draws on BioScope \cite{vincze2008bioscope}
and SFU \cite{konstantinova2012sfu} and covers five categories: epistemic phrases,
modal auxiliaries, approximators, diminishers, and conditionals. Detected cue weights
are aggregated as $s = 1 - \prod_i (1 - w_i)$, which increases monotonically with
cue count while remaining bounded. The scorer is manually validated on 50 sentences
before integration.

\paragraph{CLS Feature Injection (HedgeBERT).} The base model is BERT-base-uncased
\cite{devlin2019bert}. After encoding, the 768-dimensional CLS vector is concatenated
with the hedge scalar to produce a 769-dimensional representation, which is passed to
a re-initialized linear classification head. The model is fine-tuned on a
hedge-augmented version of the Amazon Review Dataset (1,000+ examples), with
programmatic hedge injection following the EDA framework \cite{wei2019eda}. The loss
is cross-entropy.

\paragraph{Baselines and Ablations.} The primary metric is accuracy on the hedged
and direct subsets separately for all models. Baselines include vanilla
BERT-base-uncased, DeBERTa-v3-small (zero-shot), ELECTRA-small (zero-shot), and an
LLM comparison via prompting (GPT-4o-mini). Ablations include HedgeBERT with a
random scalar in place of the hedge score (to test scorer contribution) and BERT
fine-tuned on augmented data without CLS injection (to isolate augmentation from
architecture).

\section{Evaluation Plan}

Success is defined along three dimensions. First, vanilla BERT must exhibit a
measurable accuracy gap between the hedged and direct benchmark subsets, confirming
the problem is real. Second, HedgeBERT must close or reduce that gap without
degrading direct-subset accuracy, demonstrating that CLS injection is beneficial.
Third, the scorer ablation---replacing the hedge scalar with random noise---must
degrade hedged accuracy, confirming that the scorer provides a meaningful signal. As
an additional reference point, the LLM comparison tests whether the problem persists
at scale, validating that it cannot be solved by model size alone.

\section{Timeline}

\begin{table}[h]
\centering
\small
\begin{tabular}{cl}
\toprule
\textbf{Week} & \textbf{Milestone} \\
\midrule
1--2 & Dataset curation, hedge scorer \\
3--4 & Baseline model setup, CLS injection \\
5--6 & Fine-tuning, progress presentations (April 8 \& 13) \\
7    & Evaluation \\
8    & Final report and presentation \\
\bottomrule
\end{tabular}
\end{table}

\bibliography{custom}

\end{document}
